### 2020-02-25 experiments.Rmd
Adding previous skeleton code

---
title: "Bachelor thesis"
author: "Marius Dioli"
date: "February 22, 2020"
output: pdf_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mice)
library(mvtnorm)
```

```{r}
#Setting vals for multivariate continous vars
mu = 8
sigma = 2
help(mvtnorm)
```

```{r}
#Generating binary data
n = 1000
binary_categorical_vector = round(runif(n, min = 0, max = 1))
```