### 2020-06-11 experiments.Rmd

---
title: "Bachelor thesis"
author: "Marius Dioli"
date: "February 22, 2020"
output: pdf_document
---
TODOS:
Make simulation with cross correlated variables
Incorporate FMI
Add classification and predictionY

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(mice)
library(mvtnorm)
library(MASS)
library(tidyverse) 
library(ggplot2)
```
```{r}
'?'(mvtnorm)
mu = c(1,1)
sigma = matrix(c(c(1, 0.6), c(0.6, 1)), nrow = 2, ncol = 2)
data = rmvnorm(n, mu, sigma)
data = data.frame("y_1" = data[,1], "x_1" = data[,2])
reference = lm(y_1 ~x_1, data = data)
summary(reference)

data_1_sixty = data
sixty_percent = 1/(1+exp(-(data$x_1-1.5)))
mean(sixty_percent)
for (i in 1:length(sixty_percent)){
  val = rbinom(1, 1, sixty_percent[i])
  #print(val)
  if(val == 0 ){
    data_1_sixty$y_1[i] = NA
  }
}
hist(data_1_sixty$y_1)
hist(data$y_1)
colMeans(data)
colMeans(na.omit(data_1_sixty))
mean(data_1_sixty$x_1)
data_1_sixty
```

```{r}
#Setting number of entries
n = 500
iters = 20

```

```{r}
n = 500
x_1 = rnorm(n, 1, 1)
b_1 = 8
xb = x_1*b_1+b_0
sigma_i = 10
y_1 = rnorm(n, xb, sigma_i)
df_1_full = data.frame(y_1, x_1)
reference = lm(y_1 ~x_1, data = df_1_full)
summary(reference)

```

```{r}
#Removing vals
five_percent = 1/(1+exp(-(df_1_full$x_1 +2.5)))
mean(five_percent)
ten_percent = 1/(1+exp(-(df_1_full$x_1 + 1.6)))
mean(ten_percent)
twenty_percent = 1/(1+exp(-(df_1_full$x_1+0.7)))
mean(twenty_percent)
forty_percent = 1/(1+exp(-(df_1_full$x_1-0.5)))
mean(forty_percent)
sixty_percent = 1/(1+exp(-(df_1_full$x_1-1.5)))
mean(sixty_percent)
eighty_percent = 1/(1+exp(-(df_1_full$x_1-2.65)))
mean(eighty_percent)
ninety_percent = 1/(1+exp(-(df_1_full$x_1 -3.5)))
mean(ninety_percent)
vals = 0
for(i in ninety_percent){
  if(i<0.1){
    vals = vals+1
  }
}

df_1_five = df_1_full
for (i in 1:length(five_percent)){
  val = rbinom(1, 1, five_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_five$y_1[i] = NA
  }
}

df_1_ten = df_1_full
for (i in 1:length(ten_percent)){
  val = rbinom(1, 1, ten_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_ten$y_1[i] = NA
  }
}

df_1_twenty = df_1_full
for (i in 1:length(twenty_percent)){
  val = rbinom(1, 1, twenty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_twenty$y_1[i] = NA
  }
}

df_1_forty = df_1_full
for (i in 1:length(forty_percent)){
  val = rbinom(1, 1, forty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_forty$y_1[i] = NA
  }
}

df_1_sixty = df_1_full 
for (i in 1:length(sixty_percent)){
  val = rbinom(1, 1, sixty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_sixty$y_1[i] = NA
  }
}

df_1_eighty = df_1_full
for (i in 1:length(eighty_percent)){
  val = rbinom(1, 1, eighty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_eighty$y_1[i] = NA
  }
}

df_1_ninety = df_1_full
val = rbinom(500,1, ninety_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_ninety$y_1[i] = NA
  }
}
hist(df_1_ninety$y_1)

df_1_ninety = df_1_full
for (i in 1:length(eighty_percent)){
  set.seed(i)
  val = rbinom(1, 1, ninety_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_ninety$y_1[i] = NA
  }
}
hist(df_1_ninety$y_1)
```
```{r}
df_1_five = df_1_full
val = rbinom(500,1, five_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_five$y_1[i] = NA
  }
}

df_1_ten = df_1_full
val = rbinom(500,1, ten_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_ten$y_1[i] = NA
  }
}

df_1_twenty = df_1_full
val = rbinom(500,1, twenty_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_twenty$y_1[i] = NA
  }
}

 df_1_forty = df_1_full
val = rbinom(500,1, forty_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_forty$y_1[i] = NA
  }
}

  df_1_sixty = df_1_full
val = rbinom(500,1, sixty_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_sixty$y_1[i] = NA
  }
}

  df_1_eighty = df_1_full
val = rbinom(500,1, eighty_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_eighty$y_1[i] = NA
  }
}

  df_1_ninety = df_1_full
val = rbinom(500,1, ninety_percent)
for (i in 1:length(val)){
  if(val[i] == 0 ){
    df_1_ninety$y_1[i] = NA
  }
}
'?'(dnorm)
pnorm(df_1_full$x_1, mean = 1, sd = 1)

```

```{r}
set.seed(1)

  val =pnorm(df_1_full$x_1, mean = 1, sd = 1)
df_1_five = df_1_full

for (i in 1:length(val)){
  if(val[i] < 0.5 ){
    df_1_five$y_1[i] = NA
  }
}

df_1_ten = df_1_full

for (i in 1:length(val)){
  if(val[i] < 0.1 ){
    df_1_ten$y_1[i] = NA
  }
}

df_1_twenty = df_1_full

for (i in 1:length(val)){
  if(val[i] < 0.2 ){
    df_1_twenty$y_1[i] = NA
  }
}

 df_1_forty = df_1_full

for (i in 1:length(val)){
  if(val[i] < 0.4 ){
    df_1_forty$y_1[i] = NA
  }
}

  df_1_sixty = df_1_full

for (i in 1:length(val)){
  if(val[i] < 0.6 ){

    df_1_sixty$y_1[i] = NA
  }
}
   vals = 0
  df_1_eighty = df_1_full
for (i in 1:length(val)){
  if(val[i] < 0.8 ){
    vals= vals+1
    df_1_eighty$y_1[i] = NA
  }
}

  df_1_ninety = df_1_full
for (i in 1:length(val)){
  if(val[i] < 0.9 ){
    df_1_ninety$y_1[i] = NA
  }
}
vals
hist(df_1_full$y_1)
hist(df_1_sixty$y_1)
hist(df_1_five$y_1)
hist(df_1_ten$y_1)
hist(df_1_twenty$y_1)
hist(df_1_forty$y_1)

hist(df_1_eighty$y_1)
hist(df_1_ninety$y_1)

```

```{r echo=FALSE}
#Generating data according to Turid's recommended method
#One cont normal variables
n = 500
iters = 15
imputations = 100
x_1 = rnorm(n, 1, 1)
b_1 = 1
b_0 = 0
xb = x_1*b_1 +b_0
sigma_i = 10
y_1 = rnorm(n, xb, sigma_i)
df_1_full = data.frame(y_1, x_1)
#Removing vals
five_percent = 1/(1+exp(-(df_1_full$x_1 +2.5)))
mean(five_percent)
ten_percent = 1/(1+exp(-(df_1_full$x_1 + 1.6)))
mean(ten_percent)
twenty_percent = 1/(1+exp(-(df_1_full$x_1+0.7)))
mean(twenty_percent)
forty_percent = 1/(1+exp(-(df_1_full$x_1-0.5)))
mean(forty_percent)
sixty_percent = 1/(1+exp(-(df_1_full$x_1-1.5)))
mean(sixty_percent)
eighty_percent = 1/(1+exp(-(df_1_full$x_1-2.65)))
mean(eighty_percent)
ninety_percent = 1/(1+exp(-(df_1_full$x_1 -3.5)))
mean(ninety_percent)

removers = c(five_percent, ten_percent, twenty_percent, 
             forty_percent, sixty_percent, eighty_percent, ninety_percent)

df_1_five = df_1_full
for (i in 1:length(five_percent)){
  val = rbinom(1, 1, five_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_five$y_1[i] = NA
  }
}

df_1_ten = df_1_full
for (i in 1:length(ten_percent)){
  val = rbinom(1, 1, ten_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_ten$y_1[i] = NA
  }
}

df_1_twenty = df_1_full
for (i in 1:length(twenty_percent)){
  val = rbinom(1, 1, twenty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_twenty$y_1[i] = NA
  }
}

df_1_forty = df_1_full
for (i in 1:length(forty_percent)){
  val = rbinom(1, 1, forty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_forty$y_1[i] = NA
  }
}

df_1_sixty = df_1_full 
for (i in 1:length(sixty_percent)){
  val = rbinom(1, 1, sixty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_sixty$y_1[i] = NA
  }
}

df_1_eighty = df_1_full
for (i in 1:length(eighty_percent)){
  val = rbinom(1, 1, eighty_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_eighty$y_1[i] = NA
  }
}

df_1_ninety = df_1_full
for (i in 1:length(ninety_percent)){
  val = rbinom(1, 1, ninety_percent[i])
  #print(val)
  if(val == 0 ){
    df_1_ninety$y_1[i] = NA
  }
}

#CCA
cca_1_5 = lm(y_1 ~x_1, data = df_1_five)
cca_1_10 = lm(y_1 ~ x_1, data = df_1_ten )
cca_1_20 = lm(y_1 ~ x_1, data = df_1_twenty)
cca_1_40 = lm(y_1 ~ x_1, data = df_1_forty)
cca_1_60 = lm(y_1 ~ x_1, data = df_1_sixty)
cca_1_80 = lm(y_1 ~ x_1, data = df_1_eighty)
cca_1_90 = lm(y_1 ~ x_1, data = df_1_ninety)

'?'(mice)
imp_1_5 = mice(df_1_five, m = imputations, maxit = iters, printFlag = FALSE)
imp_1_10 = mice(df_1_ten, m = imputations,maxit = iters, printFlag = FALSE)
imp_1_20 = mice(df_1_twenty,m = imputations, maxit = iters, printFlag = FALSE)
imp_1_40 = mice(df_1_forty, m = imputations,maxit = iters, printFlag = FALSE)
imp_1_60 = mice(df_1_sixty, m = imputations,maxit = iters, printFlag = FALSE)
imp_1_80 = mice(df_1_eighty,m = imputations, maxit = iters, printFlag = FALSE)
imp_1_90 = mice(df_1_ninety, m = imputations,maxit = iters, printFlag = FALSE)

fit_1_5 = with(imp_1_5, lm(y_1 ~x_1))
fit_1_10 = with(imp_1_10, lm(y_1 ~x_1))
fit_1_20 = with(imp_1_20, lm(y_1 ~x_1))
fit_1_40 = with(imp_1_40, lm(y_1 ~x_1))
fit_1_60 = with(imp_1_60, lm(y_1 ~x_1))
fit_1_80 = with(imp_1_80, lm(y_1 ~x_1))
fit_1_90 = with(imp_1_90, lm(y_1 ~x_1))

pooled_1_5 = pool(fit_1_5)
pooled_1_10 = pool(fit_1_10)
pooled_1_20 = pool(fit_1_20)
pooled_1_40 = pool(fit_1_40)
pooled_1_60 = pool(fit_1_60)
pooled_1_80 = pool(fit_1_80)
pooled_1_90 = pool(fit_1_90)

r_sq_1_5 = pool.r.squared(fit_1_5)
r_sq_1_10 = pool.r.squared(fit_1_10)
r_sq_1_20 = pool.r.squared(fit_1_20)
r_sq_1_40 = pool.r.squared(fit_1_40)
r_sq_1_60 = pool.r.squared(fit_1_60)
r_sq_1_80 = pool.r.squared(fit_1_80)
r_sq_1_90 = pool.r.squared(fit_1_90)

w = summary(pooled_1_5, conf.int=TRUE)
w
beta_ci(summary(pooled_1_5,  conf.int=TRUE))
```

```{r}
#Old
#Function for calculating confidence intervals for betas
#beta_ci = function(model_summ, lower, upper){
#  print(model_summ)
#  std = model_summ$std.error
#  estim = model_summ$estimate
#  ci_ = c()
#  for(i in 1:length(estim)){
#    print(c(estim[i] - 1.96*std[i], estim[i] + 1.96*std[i]))
#    ci_ = rbind(ci_, c(estim[i] - 1.96*std[i], estim[i] + 1.96*std[i]))
#  }
#  #print("We are at the ifs")
#  if(lower){
#    return(ci_[1,])
#  }
#  if(upper){
#    return(ci_[2,])
#  }
#  return(ci_)
#}

#New
beta_ci = function(model_summ){
  #print("We are in the function")
  w = model_summ
  ci_ = list("lower" = w$`2.5 %`[2], "upper" = w$`97.5 %`[2])
  return(ci_)
}
```

```{r}
r_sq_cca = c(summary(cca_1_5)$r.squared, summary(cca_1_10)$r.squared , summary(cca_1_20)$r.squared, summary(cca_1_40)$r.squared, summary(cca_1_60)$r.squared, summary(cca_1_80)$r.squared, summary(cca_1_90)$r.squared)
r_sq_total = c(r_sq_1_5[1], r_sq_1_10[1], r_sq_1_20[1], r_sq_1_40[1], r_sq_1_60[1], r_sq_1_80[1], r_sq_1_90[1])

beta_1_mi =c(summary(pooled_1_5)$estimate[2], summary(pooled_1_10)$estimate[2], summary(pooled_1_20)$estimate[2], summary(pooled_1_40)$estimate[2], summary(pooled_1_60)$estimate[2], summary(pooled_1_80)$estimate[2], summary(pooled_1_90)$estimate[2])

beta_1_mi_ci = rbind(beta_ci(summary(pooled_1_5,  conf.int=TRUE)), beta_ci(summary(pooled_1_10,  conf.int=TRUE)), beta_ci(summary(pooled_1_20,  conf.int=TRUE)), beta_ci(summary(pooled_1_40,  conf.int=TRUE)), beta_ci(summary(pooled_1_60,  conf.int=TRUE)), beta_ci(summary(pooled_1_80,  conf.int=TRUE)), beta_ci(summary(pooled_1_90,  conf.int=TRUE)))
beta_1_mi_ci = data.frame("lower" = unlist(beta_1_mi_ci)[1:7], "upper" =unlist(beta_1_mi_ci)[8:14])

beta_1_1_cca_ci = rbind(confint(cca_1_5)[2,], confint(cca_1_10)[2,], confint(cca_1_20)[2,], confint(cca_1_40)[2,], confint(cca_1_60)[2,], confint(cca_1_80)[2,], confint(cca_1_90)[2,])

beta_both_1_cca = rbind(cca_1_5$coefficients, cca_1_10$coefficients, cca_1_20$coefficients, cca_1_40$coefficients, cca_1_60$coefficients, cca_1_80$coefficients, cca_1_90$coefficients)
summary(cca_1_90)
miss = c(5,10, 20, 40,60,80,90)

df_1_beta = data.frame("Percent_missing" = miss, "beta_1_mi" = beta_1_1_mi, "beta_1_mi_lower" = beta_1_mi_ci$lower, "beta_1_mi_upper" = beta_1_mi_ci$upper,  "beta_0_cca" = beta_both_1_cca[,1], "beta_1_cca" = beta_both_1_cca[,2], "beta_1_cca_lower" =beta_1_1_cca_ci[,1], "beta_1_cca_upper" =beta_1_1_cca_ci[,2])
#citation("mice")

```

```{r}
#Plotting

p<- ggplot(df_1_beta, aes(x=Percent_missing, y=beta_1_cca)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=beta_1_cca_lower, ymax=beta_1_cca_upper), width=.2,
                 position=position_dodge(0.05))
t<- ggplot(df_1_beta, aes(x=Percent_missing, y=beta_1_mi)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=beta_1_mi_lower, ymax=beta_1_mi_upper), width=.2,
                 position=position_dodge(0.05))
print(p)
print(t)
df_1_beta
```

```{r}
results_5 =  data.frame("beta_0_mi" = double() , "beta_0_mi_lower" = double(), "beta_0_mi_upper" = double(), "beta_1_mi" = double(), "beta_1_mi_lower" = double(), "beta_1_mi_upper" = double(),  "beta_0_cca" = double(),  "beta_0_cca_lower"= double(), "beta_0_cca_upper" =double(), "beta_1_cca" = double(), "beta_1_cca_lower" =double(), "beta_1_cca_upper" =double())
str(results_5)
results_5 = rbind(c(1.0,1.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0), results_5)
results_5
x = c(0,0)
x = rbind(c(1,1), x)
x = rbind(c(2,2), x)
x = data.frame(x)

```

```{r}
write.table(results_5, file='/home/marius/Downloads/Test/results_5.csv', row.names=FALSE)
t = df_1_beta
t$Percent_missing = NULL
write.table(t[1,], file = "/home/marius/Downloads/Test/results_5.csv", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
summary(cca_model)$coefficients[3,4]
summary(pooled_1_20)[2,6]
x_1 = rnorm(n, 10, 1)
b_1 = 1
b_0 = 0
xb = x_1*b_1 + b_0
sigma_i = 10
y_1 = rnorm(n, xb, sigma_i)
mean(xb)
```

```{r}
  x_1 = rnorm(n, 10, 1)
  b_1 = 1
  b_0 = 0
  xb = x_1*b_1 + b_0
  sigma_i = 10
  y_1 = rnorm(n, xb, sigma_i)
  df_1_full = data.frame(y_1, x_1)

  #Removing vals
  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
  mean(five_percent)
  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
  mean(ten_percent)
  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
  mean(twenty_percent)
  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
  mean(forty_percent)
  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
  mean(sixty_percent)
  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
  mean(eighty_percent)
  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
  mean(ninety_percent)
```
```{r}
set.seed(1)
  x_1 = rnorm(n, 10, 1)
  b_1 = 1
  b_0 = 0
  xb = x_1*b_1 + b_0
  sigma_i = 10
  y_1 = rnorm(n, xb, sigma_i)
  df_1_full = data.frame(y_1, x_1)

  #Removing vals
  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
  mean(five_percent)
  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
  mean(ten_percent)
  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
  mean(twenty_percent)
  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
  mean(forty_percent)
  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
  mean(sixty_percent)
  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
  mean(eighty_percent)
  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
  mean(ninety_percent)
  df_1_five = df_1_full
  for (i in 1:length(five_percent)){
    val = rbinom(1, 1, five_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_five$y_1[i] = NA
    }
  }

  df_1_ten = df_1_full
  for (i in 1:length(ten_percent)){
    val = rbinom(1, 1, ten_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_ten$y_1[i] = NA
    }
  }

  df_1_twenty = df_1_full
  for (i in 1:length(twenty_percent)){
    val = rbinom(1, 1, twenty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_twenty$y_1[i] = NA
    }
  }

  df_1_forty = df_1_full
  for (i in 1:length(forty_percent)){
    val = rbinom(1, 1, forty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_forty$y_1[i] = NA
    }
  }

  df_1_sixty = df_1_full 
  for (i in 1:length(sixty_percent)){
    val = rbinom(1, 1, sixty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_sixty$y_1[i] = NA
    }
  }

  df_1_eighty = df_1_full
  for (i in 1:length(eighty_percent)){
    val = rbinom(1, 1, eighty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_eighty$y_1[i] = NA
    }
  }

  df_1_ninety = df_1_full
  for (i in 1:length(ninety_percent)){
    val = rbinom(1, 1, ninety_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_ninety$y_1[i] = NA
    }
  }
  colMeans(df_1_full)
  colMeans(na.omit(df_1_five))
  colMeans(na.omit(df_1_ten))
  colMeans(na.omit(df_1_twenty))
  colMeans(na.omit(df_1_forty))
  colMeans(na.omit(df_1_sixty))
  colMeans(na.omit(df_1_eighty))
  colMeans(na.omit(df_1_ninety))
```

```{r}
#Big run
n = 700
iters = 10
imputations = 20
#Empty files for appending results
results_template =  data.frame("beta_0_mi" = double() , "beta_1_mi" = double(), "beta_1_mi_lower" = double(), "beta_1_mi_upper" = double(),  "beta_0_cca" = double(),  "beta_1_cca" = double(), "beta_1_cca_lower" =double(), "beta_1_cca_upper" =double(), "p-value_mi" = double(), "p-value_cca" = double())

write.table(results_template, file='/home/marius/Downloads/Test/test_v2_5.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_10.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_20.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_40.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_60.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_80.csv', sep = ",", row.names=FALSE)
write.table(results_template, file='/home/marius/Downloads/Test/test_v2_90.csv', sep = ",", row.names=FALSE)

for(i in 1:10){
  print(i)
  #Generating data according to Turid's recommended method
  #One cont normal variables
  set.seed(i)
  x_1 = rnorm(n, 10, 1)
  b_1 = 1
  b_0 = 0
  xb = x_1*b_1 + b_0
  sigma_i = 10
  y_1 = rnorm(n, xb, sigma_i)
  df_1_full = data.frame(y_1, x_1)

  #Removing vals
  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
  mean(five_percent)
  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
  mean(ten_percent)
  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
  mean(twenty_percent)
  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
  mean(forty_percent)
  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
  mean(sixty_percent)
  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
  mean(eighty_percent)
  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
  mean(ninety_percent)

  removers = c(five_percent, ten_percent, twenty_percent, 
               forty_percent, sixty_percent, eighty_percent, ninety_percent)

  df_1_five = df_1_full
  for (i in 1:length(five_percent)){
    val = rbinom(1, 1, five_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_five$y_1[i] = NA
    }
  }

  df_1_ten = df_1_full
  for (i in 1:length(ten_percent)){
    val = rbinom(1, 1, ten_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_ten$y_1[i] = NA
    }
  }

  df_1_twenty = df_1_full
  for (i in 1:length(twenty_percent)){
    val = rbinom(1, 1, twenty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_twenty$y_1[i] = NA
    }
  }

  df_1_forty = df_1_full
  for (i in 1:length(forty_percent)){
    val = rbinom(1, 1, forty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_forty$y_1[i] = NA
    }
  }

  df_1_sixty = df_1_full 
  for (i in 1:length(sixty_percent)){
    val = rbinom(1, 1, sixty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_sixty$y_1[i] = NA
    }
  }

  df_1_eighty = df_1_full
  for (i in 1:length(eighty_percent)){
    val = rbinom(1, 1, eighty_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_eighty$y_1[i] = NA
    }
  }

  df_1_ninety = df_1_full
  for (i in 1:length(ninety_percent)){
    val = rbinom(1, 1, ninety_percent[i])
    #print(val)
    if(val == 0 ){
      df_1_ninety$y_1[i] = NA
    }
  }

  #CCA
  cca_1_5 = lm(y_1 ~x_1, data = df_1_five)
  cca_1_10 = lm(y_1 ~ x_1, data = df_1_ten )
  cca_1_20 = lm(y_1 ~ x_1, data = df_1_twenty)
  cca_1_40 = lm(y_1 ~ x_1, data = df_1_forty)
  cca_1_60 = lm(y_1 ~ x_1, data = df_1_sixty)
  cca_1_80 = lm(y_1 ~ x_1, data = df_1_eighty)
  cca_1_90 = lm(y_1 ~ x_1, data = df_1_ninety)

  imp_1_5 = mice(df_1_five, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_10 = mice(df_1_ten, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_20 = mice(df_1_twenty, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_40 = mice(df_1_forty, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_60 = mice(df_1_sixty, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_80 = mice(df_1_eighty, m = imputations, maxit = iters, printFlag = FALSE)
  imp_1_90 = mice(df_1_ninety, m = imputations, maxit = iters, printFlag = FALSE)

  fit_1_5 = with(imp_1_5, lm(y_1 ~x_1))
  fit_1_10 = with(imp_1_10, lm(y_1 ~x_1))
  fit_1_20 = with(imp_1_20, lm(y_1 ~x_1))
  fit_1_40 = with(imp_1_40, lm(y_1 ~x_1))
  fit_1_60 = with(imp_1_60, lm(y_1 ~x_1))
  fit_1_80 = with(imp_1_80, lm(y_1 ~x_1))
  fit_1_90 = with(imp_1_90, lm(y_1 ~x_1))

  pooled_1_5 = pool(fit_1_5)
  pooled_1_10 = pool(fit_1_10)
  pooled_1_20 = pool(fit_1_20)
  pooled_1_40 = pool(fit_1_40)
  pooled_1_60 = pool(fit_1_60)
  pooled_1_80 = pool(fit_1_80)
  pooled_1_90 = pool(fit_1_90)

  r_sq_1_5 = pool.r.squared(fit_1_5)
  r_sq_1_10 = pool.r.squared(fit_1_10)
  r_sq_1_20 = pool.r.squared(fit_1_20)
  r_sq_1_40 = pool.r.squared(fit_1_40)
  r_sq_1_60 = pool.r.squared(fit_1_60)
  r_sq_1_80 = pool.r.squared(fit_1_80)
  r_sq_1_90 = pool.r.squared(fit_1_90)

  #Getting p-values
  p_vals_mi = c(summary(pooled_1_5)[2,6], summary(pooled_1_10)[2,6],summary(pooled_1_20)[2,6],summary(pooled_1_40)[2,6],summary(pooled_1_60)[2,6],summary(pooled_1_80)[2,6],summary(pooled_1_90)[2,6])
  p_vals_cca = c(summary(cca_1_5)$coefficients[2,4], summary(cca_1_10)$coefficients[2,4],summary(cca_1_20)$coefficients[2,4],summary(cca_1_40)$coefficients[2,4],summary(cca_1_60)$coefficients[2,4],summary(cca_1_80)$coefficients[2,4],summary(cca_1_90)$coefficients[2,4])

  r_sq_cca = c(summary(cca_1_5)$r.squared, summary(cca_1_10)$r.squared , summary(cca_1_20)$r.squared,        summary(cca_1_40)$r.squared, summary(cca_1_60)$r.squared, summary(cca_1_80)$r.squared, summary(cca_1_90)$r.squared)
  r_sq_total = c(r_sq_1_5[1], r_sq_1_10[1], r_sq_1_20[1], r_sq_1_40[1], r_sq_1_60[1], r_sq_1_80[1], r_sq_1_90[1])

  beta_0_1_mi =c(summary(pooled_1_5)$estimate[1], summary(pooled_1_10)$estimate[1], summary(pooled_1_20)$estimate[1],   summary(pooled_1_40)$estimate[1], summary(pooled_1_60)$estimate[1], summary(pooled_1_80)$estimate[1],   summary(pooled_1_90)$estimate[1])

  beta_1_1_mi =c(summary(pooled_1_5)$estimate[2], summary(pooled_1_10)$estimate[2], summary(pooled_1_20)$estimate[2],   summary(pooled_1_40)$estimate[2], summary(pooled_1_60)$estimate[2], summary(pooled_1_80)$estimate[2],   summary(pooled_1_90)$estimate[2])

  beta_1_mi = cbind(beta_0_1_mi, beta_1_1_mi)

  beta_1_mi_ci = rbind(beta_ci(summary(pooled_1_5,  conf.int=TRUE)), beta_ci(summary(pooled_1_10,  conf.int=TRUE)), beta_ci(summary(pooled_1_20,  conf.int=TRUE)), beta_ci(summary(pooled_1_40,  conf.int=TRUE)), beta_ci(summary(pooled_1_60,  conf.int=TRUE)), beta_ci(summary(pooled_1_80,  conf.int=TRUE)), beta_ci(summary(pooled_1_90,  conf.int=TRUE)))
beta_1_mi_ci = data.frame("lower" = unlist(beta_1_mi_ci)[1:7], "upper" =unlist(beta_1_mi_ci)[8:14])

  beta_0_1_cca_ci = rbind(confint(cca_1_5)[1,], confint(cca_1_10)[1,], confint(cca_1_20)[1,], confint(cca_1_40)[1,],   confint(cca_1_60)[1,], confint(cca_1_80)[1,], confint(cca_1_90)[1,])

  beta_1_1_cca_ci = rbind(confint(cca_1_5)[2,], confint(cca_1_10)[2,], confint(cca_1_20)[2,], confint(cca_1_40)[2,],   confint(cca_1_60)[2,], confint(cca_1_80)[2,], confint(cca_1_90)[2,])
  beta_both_1_cca = rbind(cca_1_5$coefficients, cca_1_10$coefficients, cca_1_20$coefficients, cca_1_40$coefficients,   cca_1_60$coefficients, cca_1_80$coefficients, cca_1_90$coefficients)
  summary(cca_1_90)

  df_1_beta = data.frame("beta_0_mi" = beta_0_1_mi , "beta_1_mi" = beta_1_1_mi, "beta_1_mi_lower" = beta_1_mi_ci$lower, "beta_1_mi_upper" = beta_1_mi_ci$upper,  "beta_0_cca" = beta_both_1_cca[,1], "beta_1_cca" = beta_both_1_cca[,2], "beta_1_cca_lower" =beta_1_1_cca_ci[,1],   "beta_1_cca_upper" =beta_1_1_cca_ci[,2], "p-value_mi" = p_vals_mi, "p-value_cca" = p_vals_cca)
  #print(df_1_beta)

  write.table(df_1_beta[1,], file = '/home/marius/Downloads/Test/test_v2_5.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[2,], file = '/home/marius/Downloads/Test/test_v2_10.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[3,], file = '/home/marius/Downloads/Test/test_v2_20.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[4,], file = '/home/marius/Downloads/Test/test_v2_40.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[5,], file = '/home/marius/Downloads/Test/test_v2_60.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[6,], file = '/home/marius/Downloads/Test/test_v2_80.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)
  write.table(df_1_beta[7,], file = '/home/marius/Downloads/Test/test_v2_90.csv', sep = ",", append = TRUE, quote = FALSE,
    col.names = FALSE, row.names = FALSE)

}
```

```{r}
p_results_5 = read.csv('/home/marius/Downloads/Test/v2_results_5.csv')
p_results_5
```

```{r}
#Loading experiment data
results_5 = read.csv('/home/marius/Downloads/Test/test_v2_5.csv')
results_10 = read.csv('/home/marius/Downloads/Test/test_v2_10.csv')
results_20 = read.csv('/home/marius/Downloads/Test/test_v2_20.csv')
results_40 = read.csv('/home/marius/Downloads/Test/test_v2_40.csv')
results_60 = read.csv('/home/marius/Downloads/Test/test_v2_60.csv')
results_80 = read.csv('/home/marius/Downloads/Test/test_v2_80.csv')
results_90 = read.csv('/home/marius/Downloads/Test/test_v2_90.csv')

#Checking amount of times we contain the true value
tru_val = function(df){
  mi = c()
  cca = c()
  for(i in 1:length(df[,1])){
    upper_mi = df[i,4]
    lower_mi = df[i,3]
    t = FALSE
    if(1<upper_mi&&lower_mi<1){
      t = TRUE
    }
    if(t){
      mi = c(mi, as.numeric(t))
    }
    else{
      mi = c(mi, as.numeric(t))
    }

    upper_cca = df[i,8]
    lower_cca = df[i,7]
    t=FALSE
    if(1<upper_cca&&lower_cca<1){
      t = TRUE
    }
    if(t){
      cca = c(cca, as.numeric(t))
    }
    else{
      cca = c(cca, as.numeric(t))
    }

  }
  return(list("mi" = mi,"cca" = cca))
}

tru_val_5 = tru_val(results_5)
tru_val_10 = tru_val(results_10)
tru_val_20 = tru_val(results_20)
tru_val_40 = tru_val(results_40)
tru_val_60 = tru_val(results_60)
tru_val_80 = tru_val(results_80)
tru_val_90 = tru_val(results_90)

#Getting means
a=data.frame(colMeans(results_5))
r_5 = data.frame(t(a$colMeans.results_5.))
colnames(r_5) = rownames(a)

b=data.frame(colMeans(results_10))
r_10 = data.frame(t(b$colMeans.results_10.))
colnames(r_10) = rownames(b)

c=data.frame(colMeans(results_20))
r_20 = data.frame(t(c$colMeans.results_20.))
colnames(r_20) = rownames(c)

d=data.frame(colMeans(results_40))
r_40 = data.frame(t(d$colMeans.results_40.))
colnames(r_40) = rownames(d)

e=data.frame(colMeans(results_60))
r_60 = data.frame(t(e$colMeans.results_60.))
colnames(r_60) = rownames(e)

f=data.frame(colMeans(results_80))
r_80 = data.frame(t(f$colMeans.results_80.))
colnames(r_80) = rownames(f)

g=data.frame(colMeans(results_90))
r_90 = data.frame(t(g$colMeans.results_90.))
colnames(r_90) = rownames(g)

df_111 = rbind(r_5,r_10,r_20,r_40,r_60,r_80,r_90)
df_111
p<- ggplot(df_111, aes(x=c(5,10,20,40,60,80,90), y=beta_1_mi)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=beta_1_mi_lower, ymax=beta_1_mi_upper), width=.2,
                 position=position_dodge(0.05))
t<- ggplot(df_111, aes(x=c(5,10,20,40,60,80,90), y=beta_1_cca)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=beta_1_cca_lower, ymax=beta_1_cca_upper), width=.2,
                 position=position_dodge(0.05))

#Calculating bias
mi_bias = c()
cca_bias = c()
for(i in 1:length(df_111[,1])){
  bias_mi = 1-df_111[i,2] 
  bias_cca = 1-df_111[i, 6]
  mi_bias = c(mi_bias, bias_mi)
  cca_bias = c(cca_bias, bias_cca)
}
cca_bias
mi_bias
t
p

#Calculating power

#Calculating average size of CI
ci_size = function(df){
  mi = c()
  cca = c()
  for(i in 1:length(df[,1])){
    upper_mi = df[i,4]
    lower_mi = df[i,3]
    mi = c(mi, upper_mi-lower_mi)

    upper_cca = df[i,8]
    lower_cca = df[i,7]

    cca = c(cca, upper_cca-lower_cca)

  }
  return(list("mi" = mi,"cca" = cca))
}

ci_size_5 = ci_size(results_5)
ci_size_5 = data.frame(ci_size_5)
ci_size_10 = ci_size(results_10)
ci_size_10 = data.frame(ci_size_10)
ci_size_20 = ci_size(results_20)
ci_size_20 = data.frame(ci_size_20)
ci_size_40 = ci_size(results_40)
ci_size_40 = data.frame(ci_size_40)
ci_size_60 = ci_size(results_60)
ci_size_60 = data.frame(ci_size_60)
ci_size_80 = ci_size(results_80)
ci_size_80 = data.frame(ci_size_80)
ci_size_90 = ci_size(results_90)
ci_size_90 = data.frame(ci_size_90)
colMeans(data.frame(tru_val_5))
colMeans(ci_size_40)
```

```{r}

#One cont one binomial
x_1 = rnorm(10, 1, n)
x_2 = rbinom(n, 1, 0.5)

b_1 = 1
b_2 = 4

xb = matrix(c(x_1*b_1, x_2*b_2), nrow = 2, ncol = 1)
sigma_i = matrix(c(c(1,0), c(0,1)), nrow = 2, ncol = 2)

y_2 = mvrnorm(n = n, xb, sigma_i, tol = 1e-14)

```

```{r}

#One cont one binomial, 5 cont auxiliary
x_1 = rnorm(10, 1, n)
x_2 = rbinom(n, 1, 0.5)
z_1 = rnorm(10,1,n)
z_1 = rnorm(10,1,n)
z_1 = rnorm(10,1,n)
z_1 = rnorm(10,1,n)
b_1 = 1
b_2 = 0.8

```

```{r}

#Simpler version with 3 variables correlated with Y
sigma_test = matrix(c(c(1, 0.8, 0.4), c(0.8, 1, 0), c(0.4, 0, 1)), nrow = 3, ncol = 3)
eigen(sigma_test)
eigen(matrix(c(c(1, 0.8, 0.4, 0.6), c(0.8, 1, 0, 0), c(0.4, 0, 1, 0), c(0.6, 0, 0, 1)), nrow = 4, ncol = 4))
mu_test = replicate(3, 0)
simple_data = mvrnorm(n = n, mu_test, sigma_test, tol = 1e-14)
df = data.frame(simple_data)

#Playing around with simple model
play_cov = matrix(c(c(10, sqrt(10)*0.8, sqrt(10)*0.4), c(sqrt(10)*0.8, 1, 0), c(sqrt(10)*0.4, 0, 1)), nrow = 3, ncol = 3)
play_mu = c(180, 0, 0)
play = mvrnorm(n = n, play_mu, play_cov, tol = 1e-14)
df_p = data.frame(play)

```

```{r}
#Removing data

#Using the logit function to determine 
test_logit = df$X2 + df$X3
test_prob = 1/(1+exp(-(1.879+df$X2 + df$X3)))

mean(test_prob)

mar_list = c()
for (item in test_prob){
  mar_list = c(mar_list, c(rbinom(1, 100, item)))

}
mar_list
mean(mar_list)
test_prob[1]
range(length(test_prob))
#Removing values from Y
for (i in 1:length(test_prob)){
  val = rbinom(1, 1, test_prob[i])
  #print(val)
  if(val == 0 ){
    df$X1[i] = NA
  }
}

#Play
play_prob = 1/(1+exp(-(-1.5+ df_p$X2 + df_p$X3)))
mean(play_prob)
df_p_full = df_p
for (i in 1:length(play_prob)){
  val = rbinom(1, 1, play_prob[i])
  #print(val)
  if(val == 0 ){
    df_p$X1[i] = NA
  }
}
df_p
#Test
#mylogit = glm(df$X1~ df$X2 + df$X3, data = df, family = "binomial")
```

```{r}
#Multiple imputation
imputed = mice(df)
imputed$imp$X1

fitm = with(imputed, lm(X1 ~ X2 + X3))
summary(fitm)
summary(pool(fitm))

#Play
imp_p = mice(df_p)
imp_p$imp$X1

fit_p = with(imp_p, lm(X1 ~ X2 + X3))
summary(fit_p)
summary(pool(fit_p))
summary(pool(fit_p))
pool.r.squared(fit_p)

predict(pool(fit_p), data.frame(0,0))
```

```{r}
#Inference
#Test significance of regressions
#Compare vs CCA
#Test mu, sigma

#Linear regression
full_model = lm(X1 ~ X2 + X3, data = df_p_full)
summary(full_model)
full_model$coefficients[1] = 2
summary(full_model)

df_p_cca = na.omit(df_p)
cca_model = lm(X1 ~ X2 + X3, data = df_p_cca)
summary(cca_model)

#Logistic regression
```

```{r}
#Test for significance of logistic regression
```

```{r}
#Whole set of regressions with varying levels of missing data

#In the same graph compare CCA, MI, and True values.
# We compare CI, Bias, ,

#Wole set of regressions with varying levels of correlation

#In the same graph compare CCA, MI, and True values

#Whole set of regressions with categorical data as well.

```

```{r}
#Testing regression with multivariate normal
```